---
title: "Summary of Spatial Validation Simulations"
author: "Daniel J Hocking"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
df_sims <- read.csv(file = "Output/Sim_Spatial/Sim_Results.csv", header = TRUE, stringsAsFactors = FALSE)
load(file = "Output/Sim_Spatial/Sim_Spatial_Conditions.RData")
```


```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
theme_bw_journal <- function (base_family = "") {
  theme_grey(base_family = base_family) %+replace%
  theme(
  axis.text = element_text(size = rel(0.8)), axis.ticks = element_line(colour = "black"),
  legend.key = element_rect(colour = "grey80"), panel.background = element_rect(fill = "white",
  colour = NA), panel.border = element_rect(fill = NA,
  colour = "grey50"), panel.grid.major = element_line(colour = "grey90",
  size = 0.2), panel.grid.minor = element_line(colour = "grey98",
  size = 0.5), strip.background = element_rect(
  fill = "grey80",
  colour = "grey50", size = 0.2
  )
  )
  }
  
  theme_set(theme_bw_journal())
```


```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
df_sim_summary <- df_sims %>%
  dplyr::group_by(spatial, sp_mod, theta, SD_ou) %>%
  dplyr::select(-sim) %>%
    dplyr::filter(converge == TRUE) %>%
  dplyr::summarise_each(funs(mean(., na.rm = T)))
as.data.frame(df_sim_summary %>% dplyr::filter(sp_mod == 1)) 
as.data.frame(df_sim_summary %>% dplyr::filter(sp_mod == 0)) 

# ggplot(df_sim_summary, aes(SD_ou, max_N, colour = factor(theta), group = factor(theta))) + geom_smooth() #stat_smooth(aes(fill = factor(theta)))

df_converged <- df_sims %>%
  dplyr::group_by(spatial, sp_mod, theta, SD_ou) %>%
  dplyr::filter(converge == TRUE) %>%
  dplyr::filter(theta_hat < 15,
                mean_N_hat < 1000,
                rmse < 100,
                max_N < 1000) # no way converged

df_spatial <- df_converged %>%
  dplyr::filter(sp_mod == TRUE)

theta_string <- noquote(paste(theta_vec, sep = ", ", collapse = ", "))
sd_string <- noquote(paste(SD_vec, sep = ", ", collapse = ", "))
```

I ran `r max(df_sims$sim)` simulations of the spatial model for the White River in Vermont with `r nrow(family)` nodes (~330 stream reaches) for 1 year (no temporal or spatiotemporal components). For each iteration, I varied the two parameters that defind the Ornstein-Uhlenbeck process and therefore the latent spatial correlation structure in the watershed. We used all combinations of $/theta$ (`r theta_string`) and $/sigma_{ou}$ (`r sd_string`). For each combination in each iteration, I fit the model including with and without spatial correlation.


```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=10, fig.width=10}
# consider separating this out and just summarizing for the maximum number of sites or n_sites = 100 to more clearly isolate the year effects

# Make the plots
y_N <- ggplot(df_converged, aes(theta, mean_N_hat, group = interaction(theta, sp_mod))) + geom_boxplot(aes(fill = factor(sp_mod)), outlier.colour = "NA", width = 0.45) + scale_fill_manual(values=c("#F8766D", "#619CFF"), name = "Spatial", labels = c("FALSE", "TRUE")) + theme(legend.position = "right") + ylab("Mean density")  + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=2, aes(theta, mean_N, group = interaction(theta))) + geom_text(x = 5, y = 65, label = "b", size = 10) + coord_cartesian(ylim = c(0, 85)) # + ylim(c(0,500)) #+ geom_hline(yintercept = mean(df_converged$mean_N, na.rm = T), linetype = 2, size = 1, colour = "red") 

y_N_se <- ggplot(df_converged, aes(theta, N_se, group = interaction(theta, sp_mod))) + geom_boxplot(aes(fill = factor(sp_mod)), outlier.colour = "NA", width = 0.45) + scale_fill_manual(values=c("#F8766D", "#619CFF")) + theme(legend.position = "none") + ylab("SE of mean density") + geom_text(x = 5, y = 1.3, label = "c", size = 10) + coord_cartesian(ylim = c(0, 3))

y_gamma <- ggplot(df_converged, aes(theta, coef_hat, group = interaction(theta, sp_mod))) + geom_boxplot(aes(fill = factor(sp_mod)), outlier.colour = "NA", width = 0.45) + scale_fill_manual(values=c("#F8766D", "#619CFF")) + geom_hline(yintercept = df_converged$coef_true, linetype = 2, size = 0.8, colour = "darkred") + theme(legend.position = "none") + ylab(expression(paste("Fixed effect coefficient (", gamma, ")"))) + geom_text(x = 5, y = 0.72, label = "e", size = 10) + coord_cartesian(ylim = c(0.38, 0.62))

y_RMSE <- ggplot(df_converged, aes(theta, rmse, group = interaction(theta, sp_mod))) + geom_boxplot(aes(fill = factor(sp_mod)), outlier.colour = "NA", width = 0.45) + scale_fill_manual(values=c("#F8766D", "#619CFF")) + theme(legend.position = "none") + ylab("RMSE") + coord_cartesian(ylim = c(0, 80)) # + geom_text(x = 0, y = 55, label = "d", size = 10)

y_theta <- ggplot(df_spatial, aes(theta, theta_hat, group = theta)) + geom_boxplot(aes(fill = "#619CFF"), outlier.colour = "NA", width = 0.45, width = 0.45) + scale_fill_manual(values=c("#619CFF"))  + theme(legend.position="none") + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=2, aes(theta, theta, group = theta)) + ylab(expression(paste("Expected spatial decay (", theta[epsilon], ")"))) + xlab(expression(theta)) + geom_text(x = 5, y = 10.5, label = "a", size = 10) + coord_cartesian(ylim = c(0, 7)) #+ geom_text(data = means, aes(label = weight, y = weight + 0.08))
  # geom_hline(yintercept = df_spatial$theta, linetype = 2, size = 1, colour = "red") + xlab("")

# Save the legend
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

legend <- get_legend(y_N)

# Remove the legend from the box plot
y_N <- y_N + theme(legend.position="none")

# Create a blank plot
# library(cowplot)
# blank_plot <- ggplot()+geom_blank(aes(1,1)) + 
#   cowplot::theme_nothing()

# Create grid of plots
library(gridExtra)
# grid.arrange(y_theta + xlab(""), legend, 
#              y_N + xlab(""), y_N_se + xlab(""), 
#              y_RMSE + xlab(expression(paste("True value of ", theta))), y_gamma + xlab(expression(paste("True value of ", theta))), 
#              ncol = 2, 
#              nrow = 3, 
#              widths = c(2.7, 2.7), 
#              heights = c(rep(3, times=3)))

y_grid <- arrangeGrob(y_theta + xlab(""), legend, 
             y_N + xlab(""), y_N_se + xlab(""), 
             y_RMSE + xlab(expression(paste("True spatial decay rate (", theta[epsilon], ")"))), y_gamma + xlab(expression(paste("True spatial decay rate (", theta[epsilon], ")"))), 
             ncol = 2, 
             nrow = 3, 
             widths = c(2.7, 2.7), 
             heights = c(rep(3, times=3)))

grid::grid.draw(y_grid)
ggsave(file = "Output/Sim_Spatial/theta_grid_plot.png", y_grid, dpi = 300)
ggsave(file = "Output/Sim_Spatial/theta_grid_plot.pdf", y_grid, dpi = 300)

#grid.arrange(blank_plot, legend, y_N, y_N_se, y_RMSE, y_theta, y_rhot, y_sigmat, y_thet_st, y_rho_st, ncol = 2, nrow = 5, widths = c(2.7, 2.7), heights = c(0.2, rep(2.5, times=4))) # y_gamma, 

```

Varying sigma 

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE, fig.height=10, fig.width=10}
# consider separating this out and just summarizing for the maximum number of years or n_years = 20 to more clearly isolate the site effects

# Make the plots
s_N <- ggplot(df_converged, aes(SD_ou, mean_N_hat, group = interaction(SD_ou, sp_mod))) + geom_boxplot(aes(fill = factor(sp_mod)), outlier.colour = "NA", width = 0.15) + scale_fill_manual(values=c("#F8766D", "#619CFF"), name = "Spatial", labels = c("FALSE", "TRUE")) + theme(legend.position = "right") + ylab("Mean density") + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=2, aes(SD_ou, mean_N, group = interaction(SD_ou))) + coord_cartesian(ylim = c(0, 70)) # + geom_text(x = .2, y = 40, label = "b", size = 10)  # + geom_hline(yintercept = mean(df_converged$mean_N, na.rm = T), linetype = 2, size = 1, colour = "red")

s_N_se <- ggplot(df_converged, aes(SD_ou, N_se, group = interaction(SD_ou, sp_mod))) + geom_boxplot(aes(fill = factor(sp_mod)), outlier.colour = "NA", width = 0.15) + scale_fill_manual(values=c("#F8766D", "#619CFF")) + theme(legend.position = "none") + ylab("SE of mean density") + coord_cartesian(ylim = c(0, 2.6)) # + geom_text(x = .2, y = 0.35, label = "c", size = 10) 

s_gamma <- ggplot(df_converged, aes(SD_ou, coef_hat, group = interaction(SD_ou, sp_mod))) + geom_boxplot(aes(fill = factor(sp_mod)), outlier.colour = "NA", width = 0.15) + scale_fill_manual(values=c("#F8766D", "#619CFF")) + geom_hline(yintercept = df_converged$coef_true, linetype = 2, size = 0.8, colour = "darkred") + theme(legend.position = "none") + ylab(expression(paste("Fixed effect coefficient (", gamma, ")"))) + coord_cartesian(ylim = c(0.3, 0.7)) # + geom_text(x = .2, y = 0.61, label = "e", size = 10) 

s_RMSE <- ggplot(df_converged, aes(SD_ou, rmse, group = interaction(SD_ou, sp_mod))) + geom_boxplot(aes(fill = factor(sp_mod)), outlier.colour = "NA", width = 0.15) + scale_fill_manual(values=c("#F8766D", "#619CFF")) + theme(legend.position = "none") + ylab("RMSE") + coord_cartesian(ylim = c(0, 65)) # + geom_text(x = .2, y = 28, label = "d", size = 10) 

s_theta <- ggplot(df_spatial, aes(SD_ou, theta_hat, group = SD_ou)) + geom_boxplot(aes(fill = "#619CFF"), outlier.colour = "NA", width = 0.15) + scale_fill_manual(values=c("#619CFF"))  + theme(legend.position="none") + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=2, aes(SD_ou, SD_ou, group = SD_ou)) + ylab(expression(paste("Expected spatial decay (", theta[epsilon], ")"))) + coord_cartesian(ylim = c(0, 6)) #+ geom_text(x = .2, y = 70, label = "b", size = 10) 

s_sd <- ggplot(df_spatial, aes(SD_ou, SD_ou_hat, group = SD_ou)) + geom_boxplot(aes(fill = "#619CFF"), outlier.colour = "NA", width = 0.15) + scale_fill_manual(values=c("#619CFF")) + theme(legend.position="none") + xlab("") + stat_summary(fun.y=mean, colour="darkred", geom="point", shape=18, size=2, aes(SD_ou, SD_ou, group = SD_ou)) + ylab(expression(paste("Expected spatial variation (", sigma[epsilon], ")"))) + coord_cartesian(ylim = c(0, 1.5)) # + geom_text(x = .2, y = 0.8, label = "a", size = 10)  # + geom_hline(yintercept = df_spatial$theta, linetype = 2, size = 1, colour = "red")

# Save the legend
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

legend <- get_legend(s_N)

# Remove the legend from the box plot
s_N <- s_N + theme(legend.position="none")

# Create a blank plot
# library(cowplot)
# blank_plot <- ggplot()+geom_blank(aes(1,1)) + 
#   cowplot::theme_nothing()

# Create grid of plots
library(gridExtra)
# grid.arrange(s_sd + xlab(""), legend, 
#              s_N + xlab(""), s_N_se + xlab(""), 
#              s_RMSE + xlab(expression(paste("True value of ", sigma))), s_gamma + xlab(expression(paste("True value of ", sigma))), 
#              ncol = 2, 
#              nrow = 3, 
#              widths = c(2.7, 2.7), 
#              heights = c(rep(3, times=3)))

s_grid <- arrangeGrob(s_sd + xlab(""), legend, 
             s_N + xlab(""), s_N_se + xlab(""), 
             s_RMSE + xlab(expression(paste("True asymptotic spatial variation (", sigma[epsilon], ")"))), s_gamma + xlab(expression(paste("True asymptotic spatial variation (", sigma[epsilon], ")"))), 
             ncol = 2, 
             nrow = 3, 
             widths = c(2.7, 2.7), 
             heights = c(rep(3, times=3)))

grid::grid.draw(s_grid)
ggsave(file = "Output/Sim_Spatial/sigma_grid_plot.png", s_grid, dpi = 300)
ggsave(file = "Output/Sim_Spatial/sigma_grid_plot.pdf", s_grid, dpi = 300)

#grid.arrange(blank_plot, legend, y_N, y_N_se, y_RMSE, y_theta, y_rhot, y_sigmat, y_thet_st, y_rho_st, ncol = 2, nrow = 5, widths = c(2.7, 2.7), heights = c(0.2, rep(2.5, times=4))) # y_gamma, 

```

