---
title: "Adult Summary"
author: "Daniel J Hocking"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "asis")
```

```{r load}
library(TMB)
library(dplyr)
library(minqa)
library(optimx)
library(ggplot2)
source("Functions/Input_Functions.R")

adult <- readRDS(file = "Output/W_Susquehanna_Summary.Rds")
yoy <- readRDS(file = "Output/W_Susquehanna_YOY_Summary.Rds")
# load("Output/W_Susquehanna_Stock_Recruit.RData")

# load("Data/Prepared_Data_W_Susquehanna_Combined.RData")

# remove year from X_ij now so it doesn't mess with testing of temporal and temporal-spatial mdoels
#X_ij <- X_ij[ , c("length_std", "effort_std")]
# covs <- X_ij
covs <- adult$covs
X_ij <- as.matrix(dplyr::select(covs, forest, surfcoarse, temp_mean_summer_1, temp_mean_fall_1, temp_mean_winter, temp_mean_spring, prcp_mean_summer_1, prcp_mean_fall_1, prcp_mean_winter, prcp_mean_spring))
```

## Summarize Data

```{r data summary}

# number of nodes
n_nodes <- length(unique(adult$df$child_b))

# number of sites
n_sites <- length(unique(adult$df[which(!is.na(adult$df$pass_1)), "child_name"]))

# number of stream reaches
n_reaches <- n_nodes - n_sites

# number of years
n_years <- length(unique(adult$df$year))

# range of years
range_years <- range(adult$df$year)

# number of site visits
n_site_visits <- length(unique(adult$df[which(!is.na(adult$df$pass_1)), "site_visit"]))

# mean number of visits per site
mean_visits <- n_site_visits / n_sites

# range of visits per site
site_visits <- adult$df %>%
  dplyr::group_by(child_name) %>%
  dplyr::filter(!is.na(pass_1)) %>%
  dplyr::summarise(n_site_visits = n())

min_visits <- min(site_visits$n_site_visits)
max_visits <- max(site_visits$n_site_visits)

# mean number of passes per visit

# total number of fish captured by yoy and adult

# summary of covariates
cov_list <- c("AreaSqKM", "forest", "surfcoarse", "temp_mean_summer_1", "temp_mean_fall_1", "temp_mean_winter", "temp_mean_spring", "prcp_mean_summer_1", "prcp_mean_fall_1", "prcp_mean_winter", "prcp_mean_spring")

covs <- adult$df %>%
  dplyr::select(one_of(cov_list))

# summary(covs)

cov_means <- data.frame(t(dplyr::summarise_each(covs, funs(mean))))
cov_mins <- data.frame(t(dplyr::summarise_each(covs, funs(min))))
cov_maxs <- data.frame(t(dplyr::summarise_each(covs, funs(max))))
cov_medians <- data.frame(t(dplyr::summarise_each(covs, funs(median))))

cov_summary <- data.frame(cov_means, cov_mins, cov_maxs)
cov_summary <- round(cov_summary, digits = 2)
cov_summary <- data.frame(rownames(cov_means), cov_summary)
names(cov_summary) <- c("Variable", "Mean", "Min", "Max")
rownames(cov_summary) <- NULL

library(pander)
pandoc.table(cov_summary, keep.trailing.zeros = T, style = "rmarkdown")

# distances

dist_mean <- mean(adult$df$dist_b, na.rm = TRUE)
dist_min <- min(adult$df$dist_b, na.rm = TRUE)
dist_max <- max(adult$df$dist_b, na.rm = TRUE)
dist_median <- median(adult$df$dist_b, na.rm = TRUE)

df_survey_dist <- adult$df %>%
  dplyr::group_by(child_name) %>%
  dplyr::filter(!is.na(pass_1)) %>%
  dplyr::select(child_name, dist_b) %>%
  dplyr::summarise_each(funs(mean, median, min, max))

dist_survey_mean <- mean(df_survey_dist$mean, na.rm = TRUE)
dist_survey_min <- min(df_survey_dist$min, na.rm = TRUE)
dist_survey_max <- max(df_survey_dist$max, na.rm = TRUE)
dist_survey_median <- median(df_survey_dist$median, na.rm = TRUE)

```

The West Susquehanna watershed contained `r n_nodes` nodes, comprised of `r n_sites` survey sites and `r n_reaches` stream reaches. Sites were survey in a total of `r n_years` from `r min(adult$df$year, na.rm = T)` and `r max(adult$df$year, na.rm = T)`. There were a total of `r n_site_visits` site visits with a mean of `r round(mean_visits, digits = 1)` and a range of `r min_visits` to `r max_visits` visits per site. The total drainage area of the watershed was `r round(cov_summary[which(cov_summary$Variable == "AreaSqKM"), ]$Max, digits = 0)` $km^2$ and the smallest stream had a cumulative drainage area of `r round(dplyr::summarise_each(covs, funs(min))$AreaSqKM, digits = 1)`. The median drainage area was `r round(dplyr::summarise_each(covs, funs(median))$AreaSqKM, digits = 1)`. The mean distance between nodes in the network was `r round(dist_mean, digits = 3)` and ranged from `r round(dist_min, digits = 3)` to `r round(dist_max, digits = 3)` with a median of `r round(dist_median, digits = 3)` km.

## Model Comparison

```{r AIC}
pandoc.table(format(adult$aic_table, digits = 1), keep.trailing.zeros = T, style = "rmarkdown")

pandoc.table(format(yoy$aic_table, digits = 1), keep.trailing.zeros = T, style = "rmarkdown")
```

```{r coef tables}
Parameter <- c(names(adult$SD5$value))

pandoc.table(yoy$SD5$value, keep.trailing.zeros = T, style = "rmarkdown")

df_coefs <- data.frame(Parameter = c("mu",
           "sigma.mu",
           "theta",
           "rhot",
           "sigmat",
           "theta_st",
           "SD_st",
           # var_sum_sp = adult$Report5$SDinput^2 + adult$Report5$SDinput_st^2,
           # SD_st_inf = adult$Report5$SD_st_inf,
           "rho_st",
           adult$Parameters[1:10]),
           Adult = c(exp(adult$Report5$log_mean),
           adult$Report5$sigmaIID,
           NA_real_,
           adult$Report5$rhot,
           adult$Report5$sigmat,
           adult$Report5$theta_st,
           adult$Report5$SDinput_st,
           # var_sum_sp = adult$Report5$SDinput^2 + adult$Report5$SDinput_st^2,
           # SD_st_inf = adult$Report5$SD_st_inf,
           adult$Report5$rho_st,
           adult$Report5$gamma_j), 
           YOY = c(exp(yoy$Report5$log_mean),
           yoy$Report5$sigmaIID,
           NA_real_,
           yoy$Report5$rhot,
           yoy$Report5$sigmat,
           yoy$Report5$theta_st,
           yoy$Report5$SDinput_st,
           # var_sum_sp = yoy$Report5$SDinput^2 + yoy$Report5$SDinput_st^2,
           # SD_st_inf = yoy$Report5$SD_st_inf,
           yoy$Report5$rho_st,
           yoy$Report5$gamma_j[1:2],
           NA_real_,
           yoy$Report5$gamma_j[3:5],
           NA_real_,
           yoy$Report5$gamma_j[6:8]),
           stringsAsFactors = FALSE)

# cbind(adult$SD5$value, adult$SD5$sd)
# 
# cbind(yoy$SD5$value, yoy$SD5$sd)

pandoc.table(format(df_coefs, digits = 1), keep.trailing.zeros = T, style = "rmarkdown")

summary(yoy$SD5, "fixed", p.value = TRUE)
summary(yoy$SD5, "report", p.value = TRUE)
foo <- summary(yoy$SD5, "random", p.value = TRUE)
foo <- data.frame(Parameter = rownames(foo), foo, stringsAsFactors = FALSE)
bar <- dplyr::filter(foo, !grepl("detectprob_ip|extradetectrate_i|Delta_t|Deltainput_t|lognormal_overdispersed_i|SDinput_t_b|N_ip|Nu_dt", Parameter))

pandoc.table(format(summary(yoy$SD5, "report", p.value = TRUE), digits = 1), keep.trailing.zeros = T, style = "rmarkdown")

```


Compare values to where fall in relation to simulations


```{r plot effects, eval=FALSE}



n.sims=1000
simCoef <- as.data.frame(fixef(sim(glmm.M32, n.sims=n.sims)))
names(simCoef) <- names(fixef(glmm.M32))

# Plot effect of catchment forest on occurrence prob at a typical HUC10 basin # Gelman p. 44
eff.forest <- data.frame(forest.raw=seq(0,100,length.out=100)
                         , meanJulyTemp=rep(0,100)
                         , prcp=rep(0,100)
                         , surfcoarse=rep(0,100)
                         , devel_hi=rep(0,100)
                         , agriculture=rep(0,100)
                         )

eff.forest$forest <- (eff.forest$forest.raw - mean(data.fit2$forest, na.rm=T))/sd(data.fit2$forest, na.rm=T)

sim.prob.forest <- matrix(NA, nrow=nrow(eff.forest), ncol=n.sims)
for (i in 1:n.sims){
  sim.prob.forest[,i] <- exp(simCoef[i,1] + simCoef[i,"forest"]*eff.forest$forest) / (1 + exp(simCoef[i,1] + simCoef[i,"forest"]*eff.forest$forest))
}
sim.prob.forest <- as.data.frame(sim.prob.forest)

eff.forest$mean <- apply(sim.prob.forest[,1:n.sims], 1, mean)
eff.forest$lower <- apply(sim.prob.forest[,1:n.sims], 1, quantile, probs=c(0.025))
eff.forest$upper <- apply(sim.prob.forest[,1:n.sims], 1, quantile, probs=c(0.975))

ggplot(eff.forest, aes(x = forest.raw, y = mean)) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), fill="grey") +
  geom_line(colour = "black", size = 2) +
  #labs(title = "Occupancy in CT, MA, NH & NY") +
  xlab("Percent forest cover") +
  ylab("Occupancy probability") +
  theme_bw() + 
  ylim(0, 1) +
  theme(axis.text.y = element_text(size=15),
        axis.text.x = element_text(size=15),
        axis.title.x = element_text(size=17, face="bold"),
        axis.title.y = element_text(size=17, angle=90, face="bold"),
        plot.title = element_text(size=20))

forest <- seq(0, 100, by = 0.5)
stds_forest <- dplyr::filter(df_stds, var == "forest")
forest_std <- (forest - stds_forest$mean) / (stds_forest$sd)

mean <- exp(.25 + Report4b$log_mean + forest_std * coef_table[which(coef_table$Parameter == "forest"), "Estimate"])
low
high
ggplot(coef_table, aes(forest))


```

