---
title: "Adult Summary"
author: "Daniel J Hocking"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "asis", message = FALSE, warning = FALSE)
```

```{r load libraries, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(TMB)
library(dplyr)
library(tidyr)
library(minqa)
library(optimx)
library(ggplot2)
source("Functions/Input_Functions.R")
```


```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
theme_bw_journal <- function (base_family = "") {
  theme_grey(base_family = base_family) %+replace%
  theme(
  axis.text = element_text(size = rel(0.8)), axis.ticks = element_line(colour = "black"),
  legend.key = element_rect(colour = "grey80"), panel.background = element_rect(fill = "white",
  colour = NA), panel.border = element_rect(fill = NA,
  colour = "grey50"), panel.grid.major = element_line(colour = "grey90",
  size = 0.2), panel.grid.minor = element_line(colour = "grey98",
  size = 0.5), strip.background = element_rect(
  fill = "grey80",
  colour = "grey50", size = 0.2
  )
  )
  }
  
  theme_set(theme_bw_journal())
```

```{r load}
adult <- readRDS(file = "Output/W_Susquehanna_Summary.Rds")
yoy <- readRDS(file = "Output/W_Susquehanna_YOY_Summary.Rds")
# load("Output/W_Susquehanna_Stock_Recruit.RData")

# load("Data/Prepared_Data_W_Susquehanna_Combined.RData")

# remove year from X_ij now so it doesn't mess with testing of temporal and temporal-spatial mdoels
#X_ij <- X_ij[ , c("length_std", "effort_std")]
# covs <- X_ij
covs <- adult$covs
X_ij <- as.matrix(dplyr::select(covs, forest, surfcoarse, temp_mean_summer_1, temp_mean_fall_1, temp_mean_winter, temp_mean_spring, prcp_mean_summer_1, prcp_mean_fall_1, prcp_mean_winter, prcp_mean_spring))
```

## Summarize Data

```{r data summary}

# number of nodes
n_nodes <- length(unique(adult$df$child_b))

# number of sites
n_sites <- length(unique(adult$df[which(!is.na(adult$df$pass_1)), "child_name"]))

# number of stream reaches
n_reaches <- n_nodes - n_sites

# number of years
n_years <- length(unique(adult$df$year))

# range of years
range_years <- range(adult$df$year)

# number of site visits
n_site_visits <- length(unique(adult$df[which(!is.na(adult$df$pass_1)), "site_visit"]))

# mean number of visits per site
mean_visits <- n_site_visits / n_sites

# range of visits per site
site_visits <- adult$df %>%
  dplyr::group_by(child_name) %>%
  dplyr::filter(!is.na(pass_1)) %>%
  dplyr::summarise(n_site_visits = n())

min_visits <- min(site_visits$n_site_visits)
max_visits <- max(site_visits$n_site_visits)

# mean number of passes per visit

# total number of fish captured by yoy and adult

# summary of covariates
cov_list <- c("AreaSqKM", "forest", "surfcoarse", "temp_mean_summer_1", "temp_mean_fall_1", "temp_mean_winter", "temp_mean_spring", "prcp_mean_summer_1", "prcp_mean_fall_1", "prcp_mean_winter", "prcp_mean_spring")

covs <- adult$df %>%
  dplyr::select(one_of(cov_list))

# summary(covs)

cov_means <- data.frame(t(dplyr::summarise_each(covs, funs(mean))))
cov_mins <- data.frame(t(dplyr::summarise_each(covs, funs(min))))
cov_maxs <- data.frame(t(dplyr::summarise_each(covs, funs(max))))
cov_medians <- data.frame(t(dplyr::summarise_each(covs, funs(median))))

cov_summary <- data.frame(cov_means, cov_mins, cov_maxs)
cov_summary <- round(cov_summary, digits = 2)
cov_summary <- data.frame(rownames(cov_means), cov_summary)
names(cov_summary) <- c("Variable", "Mean", "Min", "Max")
rownames(cov_summary) <- NULL

library(pander)
pandoc.table(cov_summary, keep.trailing.zeros = T, style = "rmarkdown")

# distances

dist_mean <- mean(adult$df$dist_b, na.rm = TRUE)
dist_min <- min(adult$df$dist_b, na.rm = TRUE)
dist_max <- max(adult$df$dist_b, na.rm = TRUE)
dist_median <- median(adult$df$dist_b, na.rm = TRUE)

df_survey_dist <- adult$df %>%
  dplyr::group_by(child_name) %>%
  dplyr::filter(!is.na(pass_1)) %>%
  dplyr::select(child_name, dist_b) %>%
  dplyr::summarise_each(funs(mean, median, min, max))

dist_survey_mean <- mean(df_survey_dist$mean, na.rm = TRUE)
dist_survey_min <- min(df_survey_dist$min, na.rm = TRUE)
dist_survey_max <- max(df_survey_dist$max, na.rm = TRUE)
dist_survey_median <- median(df_survey_dist$median, na.rm = TRUE)

```

The West Susquehanna watershed contained `r n_nodes` nodes, comprised of `r n_sites` survey sites and `r n_reaches` stream reaches. Sites were survey in a total of `r n_years` from `r min(adult$df$year, na.rm = T)` and `r max(adult$df$year, na.rm = T)`. There were a total of `r n_site_visits` site visits with a mean of `r round(mean_visits, digits = 1)` and a range of `r min_visits` to `r max_visits` visits per site. The total drainage area of the watershed was `r round(cov_summary[which(cov_summary$Variable == "AreaSqKM"), ]$Max, digits = 0)` $km^2$ and the smallest stream had a cumulative drainage area of `r round(dplyr::summarise_each(covs, funs(min))$AreaSqKM, digits = 1)`. The median drainage area was `r round(dplyr::summarise_each(covs, funs(median))$AreaSqKM, digits = 1)`. The mean distance between nodes in the network was `r round(dist_mean, digits = 3)` and ranged from `r round(dist_min, digits = 3)` to `r round(dist_max, digits = 3)` with a median of `r round(dist_median, digits = 3)` km.

## Model Comparison

```{r AIC}
pandoc.table(format(adult$aic_table, digits = 1), keep.trailing.zeros = T, style = "rmarkdown")

pandoc.table(format(yoy$aic_table, digits = 1), keep.trailing.zeros = T, style = "rmarkdown")
```

## Adult Coefiecients

```{r coef tables}
Parameter <- c(names(adult$SD5$value))

# pandoc.table(yoy$SD5$value, keep.trailing.zeros = T, style = "rmarkdown")

df_coefs <- data.frame(Parameter = c("mu",
                                     "sigma.mu",
                                     "theta",
                                     "rhot",
                                     "sigmat",
                                     "theta_st",
                                     "SD_st",
                                     # var_sum_sp = adult$Report5$SDinput^2 + adult$Report5$SDinput_st^2,
                                     # SD_st_inf = adult$Report5$SD_st_inf,
                                     "rho_st",
                                     adult$Parameters[1:10]),
                       Adult = c(exp(adult$Report5$log_mean),
                                 adult$Report5$sigmaIID,
                                 NA_real_,
                                 adult$Report5$rhot,
                                 adult$Report5$sigmat,
                                 adult$Report5$theta_st,
                                 adult$Report5$SDinput_st,
                                 # var_sum_sp = adult$Report5$SDinput^2 + adult$Report5$SDinput_st^2,
                                 # SD_st_inf = adult$Report5$SD_st_inf,
                                 adult$Report5$rho_st,
                                 adult$Report5$gamma_j), 
                       YOY = c(exp(yoy$Report5$log_mean),
                               yoy$Report5$sigmaIID,
                               NA_real_,
                               yoy$Report5$rhot,
                               yoy$Report5$sigmat,
                               yoy$Report5$theta_st,
                               yoy$Report5$SDinput_st,
                               # var_sum_sp = yoy$Report5$SDinput^2 + yoy$Report5$SDinput_st^2,
                               # SD_st_inf = yoy$Report5$SD_st_inf,
                               yoy$Report5$rho_st,
                               yoy$Report5$gamma_j[1:2],
                               NA_real_,
                               yoy$Report5$gamma_j[3:5],
                               NA_real_,
                               yoy$Report5$gamma_j[6:8]),
                       stringsAsFactors = FALSE)

# cbind(adult$SD5$value, adult$SD5$sd)
# 
# cbind(yoy$SD5$value, yoy$SD5$sd)

# pandoc.table(format(df_coefs, digits = 1), keep.trailing.zeros = T, style = "rmarkdown")

# summary(yoy$SD5, "fixed", p.value = TRUE)
# summary(yoy$SD5, "report", p.value = TRUE)

foo <- summary(yoy$SD5, "random", p.value = FALSE)
foo <- data.frame(Parameter = rownames(foo), foo, stringsAsFactors = FALSE)
bar <- dplyr::filter(foo, !grepl("detectprob_ip|extradetectrate_i|Delta_t|Deltainput_t|lognormal_overdispersed_i|SDinput_t_b|N_ip|Nu_dt", Parameter))

summary_adult <- summary(adult$SD5, "report", p.value = FALSE)
# summary_adult <- data.frame(Parameter = c(rownames(summary_adult)[1:(length(rownames(summary_adult))-length(adult$Parameters))], adult$Parameters), summary_adult, stringsAsFactors = FALSE)
par_names <- rownames(summary_adult)
par_vec <- c(par_names[1:3], adult$Parameters, par_names[(length(par_names)-length(adult$Parameters)-2):length(par_names)])
summary_adult <- data.frame(Parameter = par_vec, summary_adult, stringsAsFactors = FALSE)

summary_yoy <- summary(yoy$SD5, "report", p.value = FALSE)
par_names <- rownames(summary_yoy)
par_vec <- c(par_names[1:3], yoy$Parameters, par_names[(length(par_names)-length(yoy$Parameters)-4):length(par_names)])
summary_yoy <- data.frame(Parameter = par_vec, summary_yoy, stringsAsFactors = FALSE)

pandoc.table(format(summary_adult, digits = 1), keep.trailing.zeros = T, style = "rmarkdown")
```

## YOY Coefficients

```{r yoy coefficients}
pandoc.table(format(summary_yoy, digits = 1), keep.trailing.zeros = T, style = "rmarkdown")
# pandoc.table(format(summary(yoy$SD5, "report", p.value = TRUE), digits = 1), keep.trailing.zeros = T, style = "rmarkdown")

summary_comb <- left_join(summary_adult, summary_yoy, by = "Parameter")
names(summary_comb) <- c("Parameter", "mu_adult", "sd_adult", "mu_yoy", "sd_yoy")
# pandoc.table(format(summary_comb, digits = 1), keep.trailing.zeros = T, style = "rmarkdown")
```

## Detection

```{r detection}
load("Data/Prepared_Data_W_Susquehanna_Combined.RData")
obs <- adult$Report5$extradetectrate_i
obs <- ifelse(obs == 1, 0, 1)

p_adult_hat <- rep(NA, 3)
for(j in 1:length(p_adult_hat)) {
  p_adult_hat[j] <- (1 - exp(-1 * adult$Report5$detectrate)) * (1 - (1 - exp(-1 * adult$Report5$detectrate))) ^ (j - 1)
}
sum(p_adult_hat)

p_yoy_hat <- rep(NA, 3)
for(j in 1:length(p_yoy_hat)) {
  p_yoy_hat[j] <- (1 - exp(-1 * yoy$Report5$detectrate)) * (1 - (1 - exp(-1 * yoy$Report5$detectrate))) ^ (j - 1)
}
sum(p_yoy_hat)

# average estimated detections including overdispersion
p_adult <- as.data.frame(adult$Report5$detectprob_ip)
names(p_adult) <- c("p1", "p2", "p3")
p_adult <- data.frame(obs, p_adult)

summary(yoy$SD5, "fixed", p.value = TRUE)
summary(yoy$SD5, "report", p.value = TRUE)
foo <- summary(yoy$SD5, "random", p.value = TRUE)
foo <- data.frame(Parameter = rownames(foo), foo, stringsAsFactors = FALSE)
bar <- dplyr::filter(foo, !grepl("detectprob_ip|extradetectrate_i|Delta_t|Deltainput_t|lognormal_overdispersed_i|SDinput_t_b|N_ip|Nu_dt", Parameter))

pandoc.table(format(summary(yoy$SD5, "report", p.value = TRUE), digits = 1), keep.trailing.zeros = T, style = "rmarkdown")

# p_adult <- p_adult %>%
#   dplyr::filter(obs == 1) %>%
#   dplyr::select(-obs)
# 
# p_adult_mean <- colMeans(p_adult) # overdispersion effect is very small on average
```

The overall mean probability of detection was `r 1 - exp(-1 * adult$Report5$detectrate)` for adults and `r 1 - exp(-1 * yoy$Report5$detectrate)` for YOY, with random variance among sites and years of `r adult$Report$extradetectionSD^2` and `r yoy$Report$extradetectionSD^2` for adults and YOY, respectively.

Compare values to where fall in relation to simulations



## Plot Adult and YOY spatio-temporal decay functions

```{r observed dist decay plot}
library(tidyr)

dist_vec <- seq(0.01, 20, by = 0.01)
Adult <- exp(-1 * adult$Report5$theta_st * dist_vec)
YOY <- exp(-1 * yoy$Report5$theta_st * dist_vec)

df_decay <- data.frame(Distance = dist_vec, Adult, YOY, stringsAsFactors = FALSE)
df_decay_long <- df_decay %>%
  tidyr::gather(Stage, rho_nu, -Distance)

g_dist <- ggplot(df_decay_long, aes(Distance, rho_nu, group = Stage)) + geom_line(aes(colour = Stage, linetype = Stage)) + xlab("Hydrologic distance (km)") + ylab(expression(paste("Spatio-temporal correlation [", hat(rho)[nu](S), "]"))) + theme_bw_journal()
g_dist

ggsave(g_dist, filename = "Output/Figures/Trout_Dist_Correlation.pdf")

g_dist <- ggplot(df_decay_long, aes(Distance, rho_nu, group = Stage)) + geom_line(aes(colour = Stage, linetype = Stage)) + xlab("Hydrologic distance (km)") + ylab(expression(paste("Spatio-temporal correlation [", hat(rho)[nu](S), "]"))) + guides(fill = NULL) +  theme_bw_journal() + theme(legend.position = "none")

```


## Plot Adult and YOY temporal decay rate

```{r observed time decay plot}
time_vec <- seq(0, 10, by = 1)
Adult <- adult$Report5$rhot ^ time_vec
YOY <- yoy$ Report5$rhot ^ time_vec

df_decay <- data.frame(Time = time_vec, Adult, YOY, stringsAsFactors = FALSE)
df_decay_long <- df_decay %>%
  tidyr::gather(Stage, rhot, -Time)

g_time <- ggplot(df_decay_long, aes(Time, rhot, group = Stage)) + geom_line(aes(colour = Stage, linetype = Stage)) + xlab("Time (years)") + ylab(expression(paste("Temporal correlation [", hat(rho)[t], "]"))) + theme_bw_journal()
g_time

ggsave(g_time, filename = "Output/Figures/Trout_Time_Correlation.pdf")

g_time <- ggplot(df_decay_long, aes(Time, rhot, group = Stage)) + geom_line(aes(colour = Stage, linetype = Stage)) + xlab("Time (years)") + ylab(expression(paste("Temporal correlation [", hat(rho)[t], "]"))) + theme_bw_journal() + theme(legend.position = "none")
g_time
```

### Combine plots

```{r combined decay plots}
library(gridExtra)
  
g_decays <- arrangeGrob(g_dist, g_time, ncol = 2)
plot(g_decays)

ggsave(g_decays, filename = "Output/Figures/Trout_Decay_Correlations.pdf")
```

## Abundances over time

```{r abundance over time, echo = FALSE}
# expected not including random overdispersion across entire watershed
lambda_t_adult <- colMeans(adult$Report5$lambda_dt) * 100
lambda_t_yoy <- colMeans(yoy$Report5$lambda_dt) * 100

# expected at sites with observed data
df_expect <- data.frame(year = t_i, obs = obs, adult = adult$Report5$N_ip[ , 1], yoy = yoy$Report5$N_ip[ , 1]) 

df_exp_summary <- df_expect %>%
  dplyr::filter(obs == 1) %>%
  dplyr::select(-obs) %>%
  dplyr::group_by(year) %>%
  dplyr::summarise_each(funs(mean))

df_exp_summary <- tidyr::gather(df_exp_summary, key = stage, value = abundance, -year)

ggplot(df_exp_summary, aes(year, abundance, colour = stage)) + geom_point() + geom_line()

ggsave(filename = "Output/Figures/Abundance_Time.pdf")
```

## Evaluate Model Fit

```{r fitted eval}
# adult
chat_adult <- tidyr::gather(as.data.frame(adult$Report5$chat_ip), key = pass, value = chat)
c_adult <- tidyr::gather(as.data.frame(c_ip), key = pass, value = count)

df_chat_adult <- bind_cols(dplyr::select(chat_adult, -pass), c_adult)
df_chat_adult <- df_chat_adult %>%
  dplyr::filter(!is.na(count))

ggplot(df_chat_adult, aes(count, chat)) + geom_point() + geom_abline(intercept = 0, slope = 1, colour = "red")
ggsave(filename = "Output/Figures/Adult_Count_Fit.pdf")

# yoy
chat_yoy <- tidyr::gather(as.data.frame(yoy$Report5$chat_ip), key = pass, value = chat)
c_yoy <- tidyr::gather(as.data.frame(c_ip), key = pass, value = count)

df_chat_yoy <- bind_cols(dplyr::select(chat_yoy, -pass), c_yoy)
df_chat_yoy <- df_chat_yoy %>%
  dplyr::filter(!is.na(count))

ggplot(df_chat_yoy, aes(count, chat)) + geom_point() + geom_abline(intercept = 0, slope = 1, colour = "red")
ggsave(filename = "Output/Figures/yoy_Count_Fit.pdf")
```


```{r plot effects, eval=FALSE}
n.sims=1000
simCoef <- as.data.frame(fixef(sim(glmm.M32, n.sims=n.sims)))
names(simCoef) <- names(fixef(glmm.M32))

# Plot effect of catchment forest on occurrence prob at a typical HUC10 basin # Gelman p. 44
eff.forest <- data.frame(forest.raw=seq(0,100,length.out=100)
                         , meanJulyTemp=rep(0,100)
                         , prcp=rep(0,100)
                         , surfcoarse=rep(0,100)
                         , devel_hi=rep(0,100)
                         , agriculture=rep(0,100)
                         )

eff.forest$forest <- (eff.forest$forest.raw - mean(data.fit2$forest, na.rm=T))/sd(data.fit2$forest, na.rm=T)

sim.prob.forest <- matrix(NA, nrow=nrow(eff.forest), ncol=n.sims)
for (i in 1:n.sims){
  sim.prob.forest[,i] <- exp(simCoef[i,1] + simCoef[i,"forest"]*eff.forest$forest) / (1 + exp(simCoef[i,1] + simCoef[i,"forest"]*eff.forest$forest))
}
sim.prob.forest <- as.data.frame(sim.prob.forest)

eff.forest$mean <- apply(sim.prob.forest[,1:n.sims], 1, mean)
eff.forest$lower <- apply(sim.prob.forest[,1:n.sims], 1, quantile, probs=c(0.025))
eff.forest$upper <- apply(sim.prob.forest[,1:n.sims], 1, quantile, probs=c(0.975))

ggplot(eff.forest, aes(x = forest.raw, y = mean)) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), fill="grey") +
  geom_line(colour = "black", size = 2) +
  #labs(title = "Occupancy in CT, MA, NH & NY") +
  xlab("Percent forest cover") +
  ylab("Occupancy probability") +
  theme_bw() + 
  ylim(0, 1) +
  theme(axis.text.y = element_text(size=15),
        axis.text.x = element_text(size=15),
        axis.title.x = element_text(size=17, face="bold"),
        axis.title.y = element_text(size=17, angle=90, face="bold"),
        plot.title = element_text(size=20))

forest <- seq(0, 100, by = 0.5)
stds_forest <- dplyr::filter(df_stds, var == "forest")
forest_std <- (forest - stds_forest$mean) / (stds_forest$sd)

mean <- exp(.25 + Report4b$log_mean + forest_std * coef_table[which(coef_table$Parameter == "forest"), "Estimate"])
low
high
ggplot(coef_table, aes(forest))


```

